Commutative character sheaves and geometric types for supercuspidal representations

### 2017-04-06 Geometric types.tex
Finished proof of Proposition 3.1

%\tableofcontents

@@

Recall from Section~\ref{sec:defs} that we refer to objects in category $\CS(G)$, defined as
in \cite{cunningham-roe:13a}*{Def. 1.1}, as linear character sheaves when $G$ is smooth but
not necessarily commutative.
Let $\CSab(G)$ denote the category of triples $(\cs{L},\mu,\beta)$ where $(\cs{L},\mu) \in \CS(G)$ and
$\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ is an isomorphism in $\CS(G\der)$.
A morphism $(\cs{L},\mu,\beta)\to (\cs{L}',\mu',\beta')$ is a morphism $\alpha : (\cs{L},\mu)\to (\cs{L}',\mu')$
in $\CS(G)$ such that $\beta = \beta' \circ \alpha\vert_{G\der}$.  We will define commutative character sheaves
as a subcategory of $\CSab(G)$, but we first relate $\CSab(G)$ to $\CS(G\ab)$.  Let $q : G \to G\ab$ be the quotient map.
\todo{Define $q\ab^* : \CS(G\ab) \to \CSab(G)$; missing definition of $\beta$ in the following commented section}
%Now suppose $q : G \to Q$ is a regular epimorphism of smooth group schemes with kernel pair $(a,p)$
%\[
%\begin{tikzcd}
%H\times G
% \arrow[shift left=2]{r}{a}
%  \arrow[shift right=2,swap]{r}{p}
%&
%G 
%\arrow{r}{q}
%& 
%Q.
%\end{tikzcd}
%\]
%Let $(\cs{L}_0, \mu_0)$ be a linear character sheaf on $Q$ and set $(\cs{L}, \mu) \ceq (q^*\cs{L}_0, (q\times q)^* \mu_0)$.
%Consider the functor
%\[
%q^* : \CS(Q) \to \CS(G)
%\]
%given on objects by $(\cs{L}_0,\mu_0) \mapsto (\cs{L}, \mu)$; see \cite{cunningham-roe:13a}*{Lem 1.4}.
%%To see that $(q^*\cs{L}_0, (q\times q)^* \mu_0)$ is indeed a linear character sheaf on $G$,
%%verify \cite{cunningham-roe:13a}, CS.3], arguing as in \cite{cunningham-roe:13a}].
%%The linear character sheaf $(\cs{L},\mu)$ on $G$ comes equipped with a canonical isomorphism $\nu : a^* \cs{L} \to p^* \cs{L}$ defined by the following diagram of isomorphisms.
%\[
%\begin{tikzcd}[column sep=40]
%a^*\cs{L} \arrow[equal]{d} \arrow[dashed]{r}{\nu} &  \arrow[equal]{d} p^*\cs{L} \\
%a^* ( q^*\cs{L}_0) \arrow{r}{q\circ a = q\circ p}
%& p^* (q^*\cs{L}_0)
%\end{tikzcd}
%\]
%Since this isomorphism satisfies \eqref{ECS1}, it follows that $(\cs{L}, \mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$.
%If $\alpha_0 : (\cs{L}_0,\mu_0) \to (\cs{L}_0',\mu_0')$ is a morphism in $\CS(Q)$, 
%then $q^*(\alpha_0) : (\cs{L},\mu) \to (\cs{L}',\mu')$ satisfies \cite{cunningham-roe:13a}*{CS4}, so $\alpha$ is a morphism in $\CS(G)$.
%These simple observations define the comparison functor
%\[
%q_H^* : \CS(Q) \to CS_H(G)
%\]
%and show that the functor $q^* : \CS(Q) \to \CS(G)$ factors according to the following commuting diagram of functors
%\begin{equation}\label{qH}
%\begin{tikzcd}
%\CS(G) &\arrow{l}[swap]{q^*} 
%\CS(Q) \arrow{dl}{q_H^*}\\
%\arrow{u}{\text{forget}} \CS_{H}(G). & 
%\end{tikzcd}
%\end{equation}
%The definition of $q_H^* : \CS(Q)\to \CS_H(G)$ will be revisited in the proof of the following result.

@@

\begin{proposition} \label{prop:CSGabab}
Suppose $G$ is a smooth group scheme.  Then the functor $q\ab^* : \CS(G\ab) \to \CSab(G)$ is an equivalence of categories.
\end{proposition}
\begin{proof}
To simplify notation below, set $H=G\der$.
Since the quotient $q : G \to G\ab$ is an $H$-torsor in the fppf topology \cite{demazure:SGA3-VIA}*{Thm. 3.2}, 
it is also an $H$-torsor in the fpqc topology.  It now follows from descent theory that the comparison functor $\Loc(G\ab) \to \Loc_H(G)$ is an equivalence; argue as in \cite{Vistoli:notes}*{Thm 4.46}, for example. \todo{Need to adapt the following proof to the definition of $q\ab^*$.}

For use below, we write $L_1 : \Loc(Q) \to \Loc_H(G)$ for the comparison functor, $R_1 : \Loc_H(G) \to \Loc(Q)$ for its adjoint, and $(\epsilon_1, \eta_1)$ for the counit and unit of the adjunction.
%
Arguing as above, we also see that pull-back along $q\times q$ determines an equivalence $L_2 : \Loc(Q\times Q) \to \Loc_{H\times H}(G\times G)$.
For use below, we write $R_2 : \Loc_{H\times H}(G\times G) \to \Loc(Q\times Q)$ for its adjoint and $(\epsilon_2, \eta_2)$ for the counit and unit of the adjunction.

We may now revisit the definition of the functor $q_H^*: \CS(Q)\to \CS_H(G)$:
on objects, $q_H^*: \CS(Q)\to \CS_H(G)$ is given by $q_H^*(\cs{L}_0,\mu_0) = (L_1(\cs{L}_0), L_2(\mu_0))$; on maps, $q_H^*: \CS(Q)\to \CS_H(G)$ is given by $q_H^*(\alpha_0) = L_1(\alpha_0)$.
Direct calculation confirms \cite{cunningham-roe:13a}*{CS1, CS2, CS3}, \eqref{E1}, \eqref{E2} and \eqref{ECS1}, and therefore that $(L_1(\cs{L}_0), L_2(\mu_0))$ is an object of $\CS_H(G)$; 
likewise, $L_1(\alpha_0)$ is a map in $\CS_H(G)$ after checking \cite{cunningham-roe:13a}*{CS4} and \eqref{E3}.
%
With this description of $q_H^*: \CS(Q)\to \CS_H(G)$ is is easy to see that it is an equivalence.
Its adjoint is given on objects by $(\cs{L},\mu,\nu) \mapsto (R_1(\cs{L},\nu), R_2(\mu))$, making use of Lemma~\ref{lem:HH}, and on morphisms by $\alpha \mapsto R_1(\alpha)$.
The adjunction is built from the adjunctions $(\epsilon_1, \eta_1)$ and $(\epsilon_2, \eta_2)$ for $(L_1,R_1)$ and $(L_2, R_2)$.
\end{proof}

>>>>

\tableofcontents

@@

Recall from Section~\ref{sec:defs} that we refer to objects in category $\CS(G)$, defined as
in \cite{cunningham-roe:13a}*{Def. 1.1}, as linear character sheaves when $G$ is smooth but
not necessarily commutative.
Let $\CSab(G)$ denote the category of triples $(\cs{L},\mu,\beta)$ where $(\cs{L},\mu) \in \CS(G)$ and
$\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ is an isomorphism in $\CS(G\der)$.
A morphism $(\cs{L},\mu,\beta)\to (\cs{L}',\mu',\beta')$ is a morphism $\alpha : (\cs{L},\mu)\to (\cs{L}',\mu')$
in $\CS(G)$ such that $\beta = \beta' \circ \alpha\vert_{G\der}$.  We will define commutative character sheaves
as a subcategory of $\CSab(G)$, but we first relate $\CSab(G)$ to $\CS(G\ab)$.  Let $q : G \to G\ab$ be the quotient map.
%\todo{Define $q\ab^* : \CS(G\ab) \to \CSab(G)$; missing definition of $\beta$ in the following commented section}
%Now suppose $q : G \to Q$ is a regular epimorphism of smooth group schemes with kernel pair $(a,p)$
%\[
%\begin{tikzcd}
%H\times G
% \arrow[shift left=2]{r}{a}
%  \arrow[shift right=2,swap]{r}{p}
%&
%G 
%\arrow{r}{q}
%& 
%Q.
%\end{tikzcd}
%\]
%Let $(\cs{L}_0, \mu_0)$ be a linear character sheaf on $Q$ and set $(\cs{L}, \mu) \ceq (q^*\cs{L}_0, (q\times q)^* \mu_0)$.
%Consider the functor
%\[
%q^* : \CS(Q) \to \CS(G)
%\]
%given on objects by $(\cs{L}_0,\mu_0) \mapsto (\cs{L}, \mu)$; see \cite{cunningham-roe:13a}*{Lem 1.4}.
%%To see that $(q^*\cs{L}_0, (q\times q)^* \mu_0)$ is indeed a linear character sheaf on $G$,
%%verify \cite{cunningham-roe:13a}, CS.3], arguing as in \cite{cunningham-roe:13a}].
%%The linear character sheaf $(\cs{L},\mu)$ on $G$ comes equipped with a canonical isomorphism $\nu : a^* \cs{L} \to p^* \cs{L}$ defined by the following diagram of isomorphisms.
%\[
%\begin{tikzcd}[column sep=40]
%a^*\cs{L} \arrow[equal]{d} \arrow[dashed]{r}{\nu} &  \arrow[equal]{d} p^*\cs{L} \\
%a^* ( q^*\cs{L}_0) \arrow{r}{q\circ a = q\circ p}
%& p^* (q^*\cs{L}_0)
%\end{tikzcd}
%\]
%Since this isomorphism satisfies \eqref{ECS1}, it follows that $(\cs{L}, \mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$.
%If $\alpha_0 : (\cs{L}_0,\mu_0) \to (\cs{L}_0',\mu_0')$ is a morphism in $\CS(Q)$, 
%then $q^*(\alpha_0) : (\cs{L},\mu) \to (\cs{L}',\mu')$ satisfies \cite{cunningham-roe:13a}*{CS4}, so $\alpha$ is a morphism in $\CS(G)$.
%These simple observations define the comparison functor
%\[
%q_H^* : \CS(Q) \to CS_H(G)
%\]
%and show that the functor $q^* : \CS(Q) \to \CS(G)$ factors according to the following commuting diagram of functors
%\begin{equation}\label{qH}
%\begin{tikzcd}
%\CS(G) &\arrow{l}[swap]{q^*} 
%\CS(Q) \arrow{dl}{q_H^*}\\
%\arrow{u}{\text{forget}} \CS_{H}(G). & 
%\end{tikzcd}
%\end{equation}
%The definition of $q_H^* : \CS(Q)\to \CS_H(G)$ will be revisited in the proof of the following result.

@@

\begin{proposition} \label{prop:CSGabab}
Suppose $G$ is a smooth group scheme.  
Then pullback along $q : G \to G\ab$ defines an equivalence $q\ab^*: \CS(G\ab) \to \CSab(G)$.
\end{proposition}
\begin{proof}
%To simplify notation below, set $H=G\der$.
Since the quotient $q : G \to G\ab$ is a $G\der$-torsor in the fppf topology \cite{demazure:SGA3-VIA}*{Thm. 3.2}, 
it is also an $H$-torsor in the fpqc topology.  
It now follows from descent theory that the comparison functor $L : \Loc(G\ab) \to \Loc_{G\der}(G)$ is an equivalence; argue as in \cite{Vistoli:notes}*{Thm 4.46}, for example. 
Observe that $L$ followed by the forgetful functor $\Loc_{G\der}(G)\to  \Loc(G)$ coincides with $q^*: \Loc(G\ab)\to \Loc(G)$ and that, for $\cs{L}\ab\in \Loc(G\ab)$, we may write $L(\cs{L}\ab) = (\cs{L}, \nu)$ where $\cs{L} = q^*(\cs{L}\ab)$ and $\nu : m^*\cs{L}  \to p^*\cs{L}$ is an isomorphism; here $m : G\times G\to G$ is  multiplication and $p : G\times  G \to G$ is projection to the second component.
%
Now set $G^2 \ceq G\times G$, so $G^2\ab =  G\ab \times G\ab$ and $G^2\der = G\der \times G\der$. Let $q_2 : G^2 \to G^2\ab$ be the abelianization quotient.  
Let $L_2 : \Loc(G^2\ab) \to \Loc_{G^2\der}(G^2)$ be the comparison functor for the quotient $q^2  : G^2 \to G^2\ab$.
Then $L_2$ is also an equivalence since $q_2$ is also a $G^2\der$-torsor.

Now consider the category $\CSab'(G)$ of triples $(\cs{L},\nu,\mu)$ where $(\cs{L},\mu)\in \CS(G)$ and $(\cs{L},\nu)\in \Loc_{G\der}(G)$; thus, $\mu : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$ and $\nu :  n^*\cs{L} \to p_n^*\cs{L}$ are isomorphisms where $n : G\der\times G\to G$ is the restriction of $m$ to $G\der\times G$ and where $p_n : G\der\times G\to G$ is projection to the second component. 
Morphism are defined in the obvious way. 
Consider the functor
\[
\begin{array}{rcl}
\CS(G\ab) &\to& \CSab'(G) \\
(\cs{L}\ab,\mu\ab) &\mapsto& (L(\cs{L}\ab), L_2(\mu\ab))
\end{array}
\]
Since both $L$ and $L_2$ are equivalences, this functor $\CS(G\ab) \to \CSab'(G)$ is an equivalence of categories.

Let $j : G\der \to G$ be the kernel of $q:G \to G\ab$ and define $j\times 1_G : G\der \to G\times G$ by $i\der(h) = (j(h),1_G)$.
For $(\cs{L},\mu,\nu)\in \CSab'(G)$, $(j\times 1_G)^*(\nu) : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ is an isomorphism.
This defines a functor
\[
\begin{array}{rcl}
\CSab'(G) &\to& \CSab(G) \\
(\cs{L},\mu,\nu) &\mapsto& (\cs{L},\mu,i\der^*(\nu)).
\end{array}
\]
Define $j\times\id_G: G\der\times G\to G\times G$ by $j\times\id_G(h,g) = (j(h),g)$.
Then for $(\cs{L},\mu,\beta) \in \CSab(G)$, define $\nu : n^*\cs{L} \to p_n^*\cs{L}$ by
\[
\begin{tikzcd}
& n^* \cs{L} \arrow{dl}{m\circ (j\times\id_G) = n} \arrow{r} & p_n^*\cs{L} &  \\ 
(j\times\id_G)^* m^* \cs{L} \arrow{dr}{(j\times\id_G)^*(\mu)} &&& (\EE)_{G\der} \boxtimes \cs{L} \arrow[equal]{ul} \\
& (j\times\id_G)^*(\cs{L}\boxtimes\cs{L}) \arrow[equal]{r} & \cs{L}\vert_{G\der}\boxtimes \cs{L} \arrow{ur}{\beta\times 1}
\end{tikzcd}
\]
Then $(\cs{L},\mu,\nu)\in \CSab'(G)$.
It now follows that $\CSab'(G) \to \CSab(G)$ is an equivalence.
\end{proof}
